---
title: "Uppsala Conflict Data"
output:
  html_document:
    code_folding: hide
    fig_width: 10
    fig_height: 8
    theme: lumen
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("conflictDat.RData")
library(dygraphs); library(dplyr); library(tidyr); library(leaflet)
library(broom); library(DT); library(ggplot2); library(ggvis)
```

## Purpose

The purpose of this document is to document the process of reshaping the **Uppsala Conflict Data**. For our particular reshaping needs we are creating a long data set that will calculate the total number of people an actor has killed and actor's side within the conflict. There are two main data sets: one containing an actor's total per year (groupTotals) and one containing an actor's total per year and their side within the conflict (groupTotalsSide).

It is important to note, that this data does not currently include civilians killed during conflict; it will, however, be included in the next version.

Here are some brief snapshots of what this data looks like:

### groupTotals

```{r}
datatable(data.frame(groupTotals))
```


### groupTotalsSide

```{r}
datatable(data.frame(groupTotalsSide))
```

## Data Summaries

We can include all manners of summaries. Just for the fun of it, let's look at what some of the drug cartels are doing.

```{r}
cartelData = groupTotals %>% 
  filter(perp == "Gulf Cartel" | 
           perp == "Juarez Cartel" | 
           perp == "Sinaloa Cartel" |
           perp == "Tijuana Cartel" |
           perp == "Los Zetas" |
           perp == "La Familia" |
           perp == "Beltrán Leyva Cartel") 

cartelData %>% 
  summarize(avgDeaths = round(mean(tot))) %>% 
  data.frame %>% 
  datatable
```


## Visualizations

### Static Visuals

```{r}
ggplot(cartelData, aes(as.factor(year), tot, color = perp)) +
  geom_point(aes(size = tot), alpha = .5) +
  scale_color_brewer(name = "Actor", palette = "Dark2") +
  scale_size_continuous(guide = FALSE) +
  labs(x = "Year", y = "Total Killed") +
  lazerhawk::theme_trueMinimal()
```


### Interactive Visuals

#### Subset

```{r, eval=FALSE}
droplevels(cartelData) %>% 
  ggvis(~as.factor(year), ~tot) %>%
  layer_points(fill = ~perp)
  lazerhawk::ggClean()
  layer_points()
```


#### Everything
```{r}
as.data.frame(groupTotals) %>%
  select(year, perp, tot) %>%
  arrange(year) %>%
  spread(perp, tot) %>%
  dygraph() %>%
  dyLegend(show = "onmouseover") %>%
  dyHighlight(highlightSeriesOpts = list(strokWidth = 7), 
              highlightSeriesBackgroundAlpha = .2,
              hideOnMouseOut = TRUE) %>%
  dyCSS("dygraphLegend.css") %>% 
  dyOptions(colors = RColorBrewer::brewer.pal(8, "Paired"))
```


#### Mapping
```{r}
leaflet(data = eventDat) %>% 
  addTiles() %>% 
  addCircles(~longitude, ~latitude, popup = ~dyad_name, radius = ~best_est)
```

## Download Data